
The material in this dissertation has appeared in the following publications. Most chapters have been slightly modified from the published articles. Nevertheless, the copyright to the original articles rests with the relevant journals.

\begin{enumerate}
\item \textit{Orthogonal matrix retrieval in cryo-electron microscopy}, Tejal Bhamre, Teng Zhang, and Amit Singer, 12th IEEE International Symposium on Biomedical Imaging, pp. 1048--1052 (2015).

\item \textit{Denoising and Covariance Estimation of Single Particle Cryo-EM Images}, Tejal Bhamre, Teng Zhang, and Amit Singer, Journal of Structural Biology, 195 (1), pp. 72-81 (2016).

\item \textit{Mahalanobis Distance for Class Averaging of Cryo-EM Image}, Tejal Bhamre, Zhizhen Zhao, and Amit Singer,  14th IEEE International Symposium on Biomedical Imaging (2017), accepted.

\item \textit{Anisotropic Twicing for Single Particle Reconstruction using Autocorrelation Analysis}, Tejal Bhamre, Teng Zhang, and Amit Singer, submitted. 
\end{enumerate}

The following public presentations were based on the materials in this dissertation:
\begin{enumerate}

\item Invited talk by Teng Zhang at Imaging and Modeling in Electron Microscopy - Recent Advances, Banff International Research Station, May 2014.

\item Poster at Semidefinite Programming and Graph Algorithms, Institute for Computational and Experimental Research in Mathematics (ICERM), February 2015.

\item Seminar at the Program in Applied and Computational Mathematics, Princeton University, February 2015.

\item Poster at the 12th IEEE International Symposium on Biomedical Imaging, April 2015.

\item Invited talk at SIAM Conference on Imaging Science, May 2016.

\item Invited talk at Workshop on Computational Methods for Cryo-EM, Simons Electron Microscopy Center, October 2016.

\end{enumerate}